---
title: "SKthink: Statistical Thinking Tools - Complete Tutorial"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SKthink: Statistical Thinking Tools - Complete Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  warning = FALSE,
  message = FALSE,
  cache = FALSE
)
```

```{r setup}
library(SKthink)
library(ggplot2)
```

# Introduction

Welcome to **SKthink**, a comprehensive R package designed to streamline your statistical analysis workflow. SKthink provides an integrated suite of tools for:

- **Session Logging**: Automatically capture all console output, commands, and plots for reproducible research
- **Data Cleaning**: Handle missing values and duplicate observations efficiently
- **Descriptive Statistics**: Generate comprehensive summary statistics and frequency distributions
- **Statistical Testing**: Perform common hypothesis tests with formatted outputs
- **Data Visualization**: Create publication-ready graphics with minimal code

This tutorial provides a complete guide to using all SKthink features in real-world analyses.

# 1. Session Logging with sknote: Capturing Your Entire Analysis

## Overview

The `sknote()` function is SKthink's flagship feature for reproducible research. Unlike traditional approaches that require manual copy-pasting, `sknote()` automatically captures **everything** that happens in your R console session:

- All commands you type
- All output (including `print()`, `cat()`, `message()`)
- Warnings and error messages
- All plots and figures
- System information (R version, OS, working directory)

This makes it invaluable for:
- Creating audit trails for regulatory compliance
- Sharing reproducible analyses with collaborators
- Documenting exploratory data analysis
- Building automated reports

## Starting a Session

Begin by calling `sknote()` with your preferred output format:

```{r eval=FALSE}
# HTML format (recommended for rich output with embedded plots)
sknote(format = "html", file_path = "my_analysis.html")

# Markdown format (great for version control and GitHub)
sknote(format = "md", file_path = "analysis_log.md")

# Plain text format (simple and universal)
sknote(format = "txt", file_path = "session_log.txt")

# Clipboard format (quick sharing without files)
sknote(format = "clipboard")
```

### Parameters Explained

| Parameter | Description | Default |
|-----------|-------------|---------|
| `format` | Output format: `"html"`, `"md"`, `"txt"`, or `"clipboard"` | Required |
| `file_path` | Path where the log file will be saved | Auto-generated |
| `save_plots` | Whether to capture plots (TRUE/FALSE) | `TRUE` |
| `plot_dir` | Directory to store plot images | `"sknote_plots"` |

## Running Your Analysis

Once `sknote()` is active, everything you do is automatically captured:

```{r eval=FALSE}
# Load data
data(mtcars)

# Basic operations are captured
head(mtcars)
summary(mtcars$mpg)

# Functions with side effects
cat("Starting analysis...\n")
message("Processing data...")

# Plots are automatically saved
plot(mtcars$wt, mtcars$mpg)
hist(mtcars$mpg, col = "lightblue")

# Even warnings and errors are logged
log(-1)  # Generates warning
```

**Key Feature**: The `sink()` function is used internally to redirect ALL console output to your log file while still displaying it in the console. This ensures nothing is missed.

## Stopping the Session

When you're done, call `sknote_stop()` to finalize your log:

```{r eval=FALSE}
sknote_stop()
```

This will:
- Stop capturing console output
- Close any open HTML tags (if HTML format)
- Save the final file
- Clean up temporary files

## Format-Specific Examples

### HTML Format (Recommended)

HTML output provides the richest experience with:
- Syntax-highlighted code blocks
- Embedded base64-encoded plots (no external files needed)
- Styled output sections
- System information header

```{r eval=FALSE}
sknote(format = "html", file_path = "report.html")

# Your analysis here
x <- rnorm(100)
mean(x)
hist(x)

sknote_stop()
# Opens report.html in browser with full formatting
```

### Markdown Format

Perfect for version control and collaboration:

```{r eval=FALSE}
sknote(format = "md", file_path = "analysis.md")

# Analysis code
data(iris)
tapply(iris$Sepal.Length, iris$Species, mean)

sknote_stop()
# Can be rendered to HTML/PDF with rmarkdown
```

### Text Format

Simple, universal format readable anywhere:

```{r eval=FALSE}
sknote(format = "txt", file_path = "log.txt")

# Simple analysis
1 + 1
print("Hello World")

sknote_stop()
# Plain text file with timestamps
```

### Clipboard Format

Quick sharing without creating files:

```{r eval=FALSE}
sknote(format = "clipboard")

# Quick calculations
2 + 2
sqrt(16)

sknote_stop()
# Everything is in your clipboard, ready to paste
```

## Advanced Tips

### Multiple Sessions

You can run multiple sknote sessions by stopping one before starting another:

```{r eval=FALSE}
# Session 1: Data cleaning
sknote(format = "html", file_path = "cleaning.html")
# ... cleaning code ...
sknote_stop()

# Session 2: Analysis
sknote(format = "html", file_path = "analysis.html")
# ... analysis code ...
sknote_stop()
```

### Plot Management

Control where plots are saved:

```{r eval=FALSE}
sknote(
  format = "html",
  save_plots = TRUE,
  plot_dir = "my_figures"
)
```

### Checking Status

To verify if sknote is running:

```{r eval=FALSE}
# Check if a session is active
exists(".sknote_env") && !is.null(SKthink:::.sknote_env$callback_id)
```

## Best Practices

1. **Start Early**: Call `sknote()` at the beginning of your analysis session
2. **Use HTML for Reports**: HTML format provides the best visual output
3. **Include System Info**: The automatic system info helps with reproducibility
4. **Version Control**: Use markdown format for Git-tracked projects
5. **Regular Backups**: Stop and restart sknote periodically for long sessions

# 2. Data Cleaning

Data cleaning is an essential step before statistical analysis. SKthink provides utilities for handling missing values and duplicate observations.

## Missing Value Report

The `sk_missing()` function generates a comprehensive report of missing values in your dataset:

```{r}
data(airquality)

sk_missing(airquality)
```

This shows which variables have missing data and the percentage of missingness, helping you decide on an appropriate imputation strategy.

## Missing Value Imputation

Use `sk_impute()` to fill in missing values using various methods:

```{r}
airquality_clean <- sk_impute(
  airquality,
  method = "median"
)

sk_missing(airquality_clean)
```

Available methods include:
- `"mean"`: Replace with column mean (numeric only)
- `"median"`: Replace with column median (numeric only)
- `"mode"`: Replace with most frequent value
- `"zero"`: Replace with zero
- `"remove"`: Remove rows with missing values

## Remove Duplicate Records

The `sk_duplicate_remove()` function identifies and removes duplicate observations:

```{r}
dup_data <- mtcars[c(1,1,2,3,3,3), ]

clean_data <- sk_duplicate_remove(
  dup_data
)

nrow(dup_data)

nrow(clean_data)
```

After cleaning, the dataset becomes ready for analysis.

# 3. Descriptive Statistics and Exploratory Data Analysis

Exploratory analysis helps users understand distributions, trends, and potential anomalies in the data before formal hypothesis testing.

## Summary Statistics

The `sk_stat()` function provides comprehensive descriptive statistics for all numeric variables:

```{r}
sk_stat(mtcars)
```

This includes count, mean, standard deviation, median, min, max, quartiles, and skewness for each numeric column.

## Frequency Distribution

For categorical or discrete variables, use `sk_frequency()`:

```{r eval=FALSE}
sk_frequency(mtcars$cyl)
```

This shows counts, percentages, and cumulative percentages for each unique value.

## Group-wise Summary Statistics

Compare statistics across groups with `sk_group_stat()`:

```{r eval=FALSE}
sk_group_stat(
  data = mtcars,
  group = "cyl",
  variable = "mpg"
)
```

This generates summary statistics (mean, sd, min, max, etc.) for each level of the grouping variable, perfect for exploratory comparisons before ANOVA or t-tests.

Descriptive analysis provides an overview of the dataset before formal hypothesis testing.

# 4. Statistical Analysis

SKthink includes several statistical testing functions that provide formatted, publication-ready outputs for common analytical tasks.

## Shapiro-Wilk Normality Test

Test whether a variable follows a normal distribution:

```{r}
sk_shapiro_test(
  mtcars$mpg
)
```

The output includes the test statistic (W), p-value, and interpretation. A non-significant p-value (>0.05) suggests the data is normally distributed.

## Independent Sample T-Test

Compare means between two independent groups:

```{r}
auto_mpg <- mtcars$mpg[
  mtcars$am == 0
]

manual_mpg <- mtcars$mpg[
  mtcars$am == 1
]

sk_t_test_two(
  auto_mpg,
  manual_mpg
)
```

Results include t-statistic, degrees of freedom, p-value, confidence interval, and effect size (Cohen's d).

## Analysis of Variance (ANOVA)

Compare means across three or more groups:

```{r}
mtcars$cyl <- factor(
  mtcars$cyl
)

sk_anova(
  hp ~ cyl,
  data = mtcars
)
```

The function provides the ANOVA table with F-statistic, p-value, and post-hoc pairwise comparisons if the overall test is significant.

## Correlation Analysis

Assess the relationship between two continuous variables:

```{r}
sk_correlation_test(
  mtcars$wt,
  mtcars$mpg,
  method = "pearson"
)
```

Available methods:
- `"pearson"`: Linear correlation (default, requires normality)
- `"spearman"`: Rank-based correlation (non-parametric)
- `"kendall"`: Another rank-based method (good for small samples)

Output includes correlation coefficient (r), p-value, confidence interval, and interpretation guidelines.

These functions provide formatted outputs suitable for reports and publications, with automatic assumption checking where applicable.

# 5. Data Visualization

Visualization is an important component of statistical communication. SKthink provides wrappers around ggplot2 for generating publication-ready graphics with minimal code. All plotting functions use the consistent `sk_theme()` for professional styling.

## Scatter Plot

Visualize relationships between two continuous variables:

```{r eval=FALSE}
sk_scatter(
  data = mtcars,
  x = "wt",
  y = "mpg",
  color = "cyl",
  add_regression = TRUE,
  title = "Weight vs MPG",
  xlab = "Weight (1000 lbs)",
  ylab = "Miles Per Gallon"
)
```

Key parameters:
- `color`: Color points by a categorical variable
- `add_regression`: Add a regression line (TRUE/FALSE)
- `size`: Control point size
- `alpha`: Control transparency

## Histogram

Display the distribution of a single continuous variable:

```{r eval=FALSE}
sk_histogram(
  data = mtcars,
  x = "mpg",
  add_density = TRUE,
  bins = 10,
  title = "Distribution of MPG",
  fill_color = "steelblue"
)
```

Key parameters:
- `add_density`: Overlay a density curve (TRUE/FALSE)
- `bins`: Number of histogram bins
- `fill_color`: Bar fill color

## Box Plot

Compare distributions across categories:

```{r eval=FALSE}
sk_boxplot(
  data = mtcars,
  x = "cyl",
  y = "mpg",
  fill = "cyl",
  add_points = TRUE,
  title = "MPG by Cylinder Group",
  palette = "Set2"
)
```

Key parameters:
- `fill`: Fill boxes by a variable
- `add_points`: Overlay individual data points (TRUE/FALSE)
- `palette`: Color palette name (RColorBrewer)

## Bar Plot

Display categorical summaries:

```{r eval=FALSE}
cyl_summary <- aggregate(
  mpg ~ cyl,
  data = mtcars,
  mean
)

sk_bar(
  data = cyl_summary,
  x = "cyl",
  y = "mpg",
  title = "Average MPG by Cylinder",
  fill_color = "coral",
  add_labels = TRUE
)
```

Key parameters:
- `fill_color`: Bar fill color
- `add_labels`: Add value labels on bars (TRUE/FALSE)
- `orientation`: "vertical" or "horizontal"

## Line Plot

Show trends over time or ordered categories:

```{r eval=FALSE}
sk_line(
  data = mtcars,
  x = "carb",
  y = "mpg",
  title = "MPG by Carburetor Count",
  color = "darkgreen",
  add_points = TRUE
)
```

Key parameters:
- `color`: Line color
- `add_points`: Add points at data values (TRUE/FALSE)
- `line_type`: Line type (solid, dashed, dotted)

All plotting functions return ggplot objects, so you can further customize them using standard ggplot2 syntax:

```{r eval=FALSE}
p <- sk_scatter(mtcars, "wt", "mpg")
p + theme_minimal() + labs(caption = "Source: Motor Trend")
```

These plotting functions reduce boilerplate code and create informative visualizations quickly while maintaining consistency across all your figures.

# 6. Complete Workflow Example: End-to-End Analysis

The following example demonstrates a complete SKthink workflow from data cleaning to analysis and reporting, with full session logging for reproducibility.

## Scenario: Air Quality Analysis

Let's analyze the built-in `airquality` dataset, handling missing values and exploring relationships between variables.

```{r eval=FALSE}
# Step 1: Start session logging (captures everything!)
sknote(
  format = "html",
  file_path = "airquality_analysis.html",
  save_plots = TRUE
)

# Step 2: Load and inspect data
data(airquality)
head(airquality)

# Step 3: Check for missing values
sk_missing(airquality)

# Step 4: Impute missing values with median
airquality_clean <- sk_impute(
  airquality,
  method = "median"
)

# Verify imputation worked
sk_missing(airquality_clean)

# Step 5: Descriptive statistics
sk_stat(airquality_clean)

# Step 6: Visualize distributions
sk_histogram(
  data = airquality_clean,
  x = "Ozone",
  add_density = TRUE,
  title = "Ozone Distribution"
)

sk_histogram(
  data = airquality_clean,
  x = "Solar.R",
  title = "Solar Radiation Distribution"
)

# Step 7: Explore relationships
sk_scatter(
  data = airquality_clean,
  x = "Temp",
  y = "Ozone",
  add_regression = TRUE,
  title = "Temperature vs Ozone"
)

sk_boxplot(
  data = airquality_clean,
  x = "Month",
  y = "Ozone",
  fill = "Month",
  title = "Ozone by Month"
)

# Step 8: Statistical testing
# Test if Ozone differs by month
airquality_clean$Month <- factor(airquality_clean$Month)
sk_anova(Ozone ~ Month, data = airquality_clean)

# Correlation between Temp and Ozone
sk_correlation_test(
  airquality_clean$Temp,
  airquality_clean$Ozone,
  method = "pearson"
)

# Step 9: Stop logging and save report
sknote_stop()

# The HTML file 'airquality_analysis.html' now contains:
# - All commands executed
# - All output and results
# - All plots (embedded as base64)
# - System information
# - Timestamps
```

## What You Get

After running this workflow, you'll have:

1. **HTML Report** (`airquality_analysis.html`): A complete, self-contained report with all analysis steps, outputs, and visualizations
2. **Plot Directory** (`sknote_plots/`): Individual PNG files of all plots generated during the session
3. **Reproducible Record**: Anyone can re-run your analysis and get identical results

## Tips for Real Projects

- **Start logging immediately**: Call `sknote()` before any analysis code
- **Use meaningful filenames**: Include project name and date in log files
- **Save plots externally**: Set `save_plots = TRUE` for high-resolution figures
- **Combine with R Markdown**: Use sknote for exploration, then create final reports with R Markdown

This workflow demonstrates how SKthink can be used in real-world projects while maintaining reproducibility and professional documentation standards.

# Conclusion

SKthink provides an integrated environment for:

- **Session Logging and Reproducibility**: Automatic capture of all console output, commands, and plots with `sknote()`
- **Data Cleaning and Preprocessing**: Handle missing values and duplicates efficiently
- **Descriptive Statistics**: Comprehensive summary statistics and group-wise comparisons
- **Statistical Hypothesis Testing**: Formatted outputs for common tests (t-test, ANOVA, correlation)
- **Data Visualization**: Publication-ready graphics with consistent styling

## Why Choose SKthink?

1. **Time-Saving**: Reduce boilerplate code for common statistical tasks
2. **Reproducibility**: Built-in session logging ensures nothing is lost
3. **Professional Output**: Formatted tables and publication-ready plots
4. **Beginner-Friendly**: Simple function calls for complex operations
5. **Extensible**: All functions return standard R objects for further customization

By combining these tools into a single package, SKthink helps analysts focus on statistical thinking rather than repetitive coding tasks. Whether you're a student learning statistics, a researcher conducting analyses, or a data scientist building reports, SKthink streamlines your workflow from data import to final report.

## Getting Help

- Browse function documentation: `?sknote`, `?sk_stat`, etc.
- View this vignette: `vignette("skthink-tutorial")`
- Report issues: [GitHub Issues](https://github.com/yourusername/SKthink/issues)
- Contributing: We welcome contributions! See CONTRIBUTING.md

Happy analyzing with SKthink! 📊🔍
````
