````markdown
---
title: "SKthink: Statistical Thinking Tools"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SKthink: Statistical Thinking Tools}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5,
  warning = FALSE,
  message = FALSE
)
```

```{r setup}
library(SKthink)
library(ggplot2)
```

# Introduction

Welcome to **SKthink**, a comprehensive R package for session logging, data cleaning, statistical analysis, and data visualization. The package is designed to simplify common statistical workflows while improving reproducibility and reporting.

This vignette demonstrates a complete workflow using SKthink.

# 1. Session Logging and Reproducibility

One of the key features of SKthink is session logging. It allows users to save their analysis workflow and create reproducible records of their work.

````rs
## Start a Logging Session

```{r eval=FALSE}
sknote(
  format = "html",
  file_path = "analysis_log.html"
)
```

## Run Your Analysis

```{r eval=FALSE}
x <- 1:10

mean(x)

summary(x)
```

## Stop and Save the Session

```{r eval=FALSE}
sknote_stop()
```

```


## Check Session Status

```{r eval=FALSE}
sk_note_status()
```

The session logging module helps users maintain reproducible analysis records and share their workflow with collaborators.

# 2. Data Cleaning

Data cleaning is an essential step before statistical analysis. SKthink provides utilities for handling missing values and duplicate observations.

## Missing Value Report

```{r}
data(airquality)

sk_missing(airquality)
```

## Missing Value Imputation

```{r}
airquality_clean <- sk_impute(
  airquality,
  method = "median"
)

sk_missing(airquality_clean)
```

## Remove Duplicate Records

```{r}
dup_data <- mtcars[c(1,1,2,3,3,3), ]

clean_data <- sk_duplicate_remove(
  dup_data
)

nrow(dup_data)

nrow(clean_data)
```

After cleaning, the dataset becomes ready for analysis.

# 3. Descriptive Statistics and Exploratory Data Analysis

Exploratory analysis helps users understand distributions, trends, and potential anomalies in the data.

## Summary Statistics

```{r}
sk_stat(mtcars)
```

## Frequency Distribution

```{r eval=FALSE}
sk_frequency(mtcars$cyl)
```

## Group-wise Summary Statistics

```{r eval=FALSE}
sk_group_stat(
  data = mtcars,
  group = "cyl",
  variable = "mpg"
)
```

Descriptive analysis provides an overview of the dataset before formal hypothesis testing.

# 4. Statistical Analysis

SKthink includes several statistical testing functions for common analytical tasks.

## Shapiro-Wilk Normality Test

```{r}
sk_shapiro_test(
  mtcars$mpg
)
```

## Independent Sample T-Test

```{r}
auto_mpg <- mtcars$mpg[
  mtcars$am == 0
]

manual_mpg <- mtcars$mpg[
  mtcars$am == 1
]

sk_t_test_two(
  auto_mpg,
  manual_mpg
)
```

## Analysis of Variance (ANOVA)

```{r}
mtcars$cyl <- factor(
  mtcars$cyl
)

sk_anova(
  hp ~ cyl,
  data = mtcars
)
```

## Correlation Analysis

```{r}
sk_correlation_test(
  mtcars$wt,
  mtcars$mpg,
  method = "pearson"
)
```

These functions provide formatted outputs suitable for reports and publications.

# 5. Data Visualization

Visualization is an important component of statistical communication. SKthink provides wrappers around ggplot2 for generating publication-ready graphics.

## Scatter Plot

```{r eval=FALSE}
sk_scatter(
  data = mtcars,
  x = "wt",
  y = "mpg",
  color = "cyl",
  add_regression = TRUE,
  title = "Weight vs MPG"
)
```

## Histogram

```{r eval=FALSE}
sk_histogram(
  data = mtcars,
  x = "mpg",
  add_density = TRUE,
  title = "Distribution of MPG"
)
```

## Box Plot

```{r eval=FALSE}
sk_boxplot(
  data = mtcars,
  x = "cyl",
  y = "mpg",
  fill = "cyl",
  add_points = TRUE,
  title = "MPG by Cylinder Group"
)
```

## Bar Plot

```{r eval=FALSE}
cyl_summary <- aggregate(
  mpg ~ cyl,
  data = mtcars,
  mean
)

sk_bar(
  data = cyl_summary,
  x = "cyl",
  y = "mpg",
  title = "Average MPG by Cylinder"
)
```

These plotting functions reduce boilerplate code and create informative visualizations quickly.

# 6. Complete Workflow Example

The following example demonstrates a complete SKthink workflow from data cleaning to analysis and reporting.

```{r eval=FALSE}
sk_note()

data(airquality)

airquality_clean <- sk_impute(
  airquality,
  method = "median"
)

sk_stat(
  airquality_clean
)

sk_histogram(
  data = airquality_clean,
  x = "Ozone"
)

sk_note_close(
  "complete_analysis.txt"
)
```

This workflow demonstrates how SKthink can be used in real-world projects while maintaining reproducibility.

# Conclusion

SKthink provides an integrated environment for:

- Session logging and reproducibility
- Data cleaning and preprocessing
- Descriptive statistics
- Statistical hypothesis testing
- Data visualization

By combining these tools into a single package, SKthink helps analysts focus on statistical thinking rather than repetitive coding tasks.
````
